---
title: "Final Project Presentation"
author: "Michael Lozada and Sophia de los Reyes"
date: "12/6/2021"
output: 
  ioslides_presentation:
      widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(geojsonio)
library(htmltools)
library(leaflet)
library(caret)
library(rpart)
library(rpart.plot)
library(openintro)
library(tidymodels)
library(rsample)
library(rminer)

detach("package:dials", unload = TRUE)
```
## Background 

- Cycles of Public Action and Private Interest 
- Movement towards Deinstitutionalization and Reprivatization
- Debate: Are for-profit healthcare facilities beneficial or detrimental to its patients? 
  - Improved patient care and reduced prices
  - Quality decreases, accessibility and availability decreases 

## Research Question

- Do the treatments offered at a mental health facility predict the type of ownership of the facility: private for-profit, private non-profit, and public? 

## Methods: Finding the Data 

- National Mental Health Services Survey (2019, most recent)
- Chose Ownership 
- Selected Treatments: Psychotherapy, Emergency, Substance Abuse, PTSD, Eating Disorder, Alzheimer's or Dementia, Electroconvulsive

## Methods: Tidying Data for Visualizations 

- Recoded variables and observation cells 
- Pivot for visualizations 
  - Bar plots 
  
## Methods: Classification Tree 

- Classification Tree to predict ownership from treatments offered 
- Training and Testing Data, Pruning, Confusion Matrix



## Selected Treatments and their Availability in Facilities Across the US
```{r test slide code, echo=FALSE, include=FALSE}
PUF_new <- read_csv("~/RStudio/Final Project/Facilities Data/Sophia PUF_new.csv")

test1 <- factor(PUF_new$treatment, level = c("Psychotherapy", "Substance Abuse", "PTSD", "Emergency", "Eating Disorder", "Alzheimer's or Dementia", "Electroconvulsive"))

treat1 <- ggplot(PUF_new, aes(x = test1, fill = Availability)) + 
  geom_bar() + 
  labs(x = "Treatment",
       y = "Frequency",
       color = "Availability") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.60, hjust = 0.5)) 


```

```{r, treatment1, echo=FALSE}
treat1
```

## Faceted by Ownership
```{r, treatment2 code, echo=FALSE, include=FALSE}
  
treat2 <- ggplot(PUF_new, aes(x = test1, fill = Availability)) + 
  geom_bar() + 
  facet_grid(Ownership ~ .) +
  labs(x = "Treatment",
       y = "Frequency",
       Legend = "Availability") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.60, hjust = 0.5))

```
```{r, treatment2, echo=FALSE}
treat2
```

## Faceted by Ownership (%)
```{r, treatment3 code, echo=FALSE, include=FALSE}
treat3 <- ggplot(PUF_new, aes(x = test1, fill = Availability)) + 
  geom_bar(position = "fill") + 
  facet_grid(Ownership ~ .) +
  labs(x = "Treatment",
       y = "Frequency (Percent)",
       Legend = "Availability") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.60, hjust = 0.5))
```
```{r, treatment3, echo=FALSE}
treat3
```

## Results: Treatment Visualization

- Visualizations
  - Psychotherapy, substance abuse, and emergency treatments provided more in public institutions 
  - Eating disorder treatments provided more in for-profit institutions


## Payment Types and their Availability in Facilities Across the US
```{r, funds code, echo=FALSE, include=FALSE}
PUF_new_funds <- read_csv("~/RStudio/Final Project/Facilities Data/Sohpia PUF_new_funds.csv")

test2 <- factor(PUF_new_funds$payment, level = c("Sliding Fee", "Treatment No or Minimal Pay", "Accepts Cash or Self-Payment", "Accepts Private Health Insurance", "Accepts Medicare", "Accepts Medicaid", "Accepts State-Financed Insurance", "Accepts State MH Agency Funds", "Accepts State Welfare", "Accepts State Juvenile Agency Funds", "Accepts State Education Agency Funds", "Accepts Other State Funds", "Accepts Local Government Funds", "Accepts Community Service Block Grants", "Accepts Community MH Block Grants", "Accepts Federal Military Insurance", "Accepts U.S. Dept. of Veterans Affairs", "Accepts IHS/Tribal/Urban Funds", "Accepts Other Sources"))

funds1 <- ggplot(PUF_new_funds, aes(x = test2, fill = acceptance)) + 
  geom_bar() + 
  labs(
       x = "Payment",
       y = "Frequency",
       color = "Acceptance") +
  theme(axis.text.x = element_text(angle = 90))
  
```

```{r, funds1, echo=FALSE}
funds1
```


## Faceted by Ownership
```{r, funds2 code, echo=FALSE, include=FALSE}
funds2 <- ggplot(PUF_new_funds, aes(x = test2, fill = acceptance)) + 
  geom_bar() + 
  facet_grid(Ownership ~ .) +
  labs(
       x = "Payment",
       y = "Frequency",
       Legend = "Acceptance") +
  theme(axis.text.x = element_text(angle = 90))
```
```{r, funds2, echo=FALSE}
funds2
```

## Faceted by Ownership (%)
```{r, funds3 code, echo=FALSE, include=FALSE}
funds3 <- ggplot(PUF_new_funds, aes(x = test2, fill = acceptance)) + 
  geom_bar(position = "fill") + 
  facet_grid(Ownership ~ .) +
  labs(
       x = "Payment",
       y = "Frequency",
       Legend = "Acceptance") +
  theme(axis.text.x = element_text(angle = 90))
```
```{r, funds3, echo=FALSE}
funds3
```

## Results: Funds Visualization

- Highest accepted payment types are ones you pay for:
  - Cash 
  - Personal Health Insurance
  - Medicare/Medicaid (taxes)
- Other state provided funds are less accepted
  - Needs-based



## Classification Tree (Decision Tree)
```{r, class tree code, echo=FALSE, include=FALSE}
facilities_final <- read_csv("~/RStudio/Final Project/Facilities Data/final.csv")
set.seed(1116)
facilities_final1 <- facilities_final %>% mutate(Psychotherapy = as.factor(facilities_final$Psychotherapy)) %>% mutate(Electroconvulsive = as.factor(facilities_final$Electroconvulsive)) %>% mutate(`Substance Abuse` = as.factor(facilities_final$`Substance Abuse`)) %>% mutate(Emergency = as.factor(facilities_final$Emergency)) %>% mutate(`Alzheimer's or Dementia` = as.factor(facilities_final$`Alzheimer's or Dementia`)) %>% mutate(`Eating Disorder` = as.factor(facilities_final$`Eating Disorder`)) %>% mutate(PTSD = as.factor(facilities_final$PTSD)) %>% mutate(Ownership = as.factor(facilities_final$Ownership))
facilities_split <- initial_split(facilities_final1, prop = 0.80)
train_data <- training(facilities_split)
test_data <- testing(facilities_split)

options(repr.plot.width = 6, repr.plot.height = 5)
facilities_tree_train <- rpart(Ownership ~ Psychotherapy + Electroconvulsive + `Substance Abuse` + Emergency + `Alzheimer's or Dementia` + `Eating Disorder` + PTSD, cp = 0.001, maxdepth = 7, minbucket = 5, method = "class", data = train_data)

facilities_tree_train_prune <- prune(facilities_tree_train, cp = 0.002)
```

```{r, class tree, echo=FALSE}
prp(facilities_tree_train, type=1, extra=104, space = 4, split.cex = 1.5, nn.border.col = 1)
```


## Pruned Classification Tree
```{r, pruned class tree, echo=FALSE}
prp(facilities_tree_train_prune, type=1, extra=104, space = 4, split.cex = 1.5, nn.border.col = 1)
```

## Confusion Matrix pt. 1
```{r, confusion matrix code, echo=FALSE, include=FALSE}
fit.control <- trainControl(method = "repeatedcv", number = 5, repeats = 10)
set.seed(1116)
facilities_fit <- train(Ownership ~ Psychotherapy + Electroconvulsive + `Substance Abuse` + Emergency + `Alzheimer's or Dementia` + `Eating Disorder` + PTSD, data = train_data, method = "multinom", trControl = fit.control, trace = FALSE)

predict_facilities <- predict(facilities_fit, newdata = test_data)
str(predict_facilities)
prob_facilities <- predict(facilities_fit, newdata = test_data, type = "prob")
head(prob_facilities)
```

```{r, confusion matrix, echo=FALSE}
confusionMatrix(data = predict_facilities, test_data$Ownership)
```

## Confusion Matrix pt. 2

- Statistics by Class:

- Sensitivity
  - Private For-Profit: 0.027027
  - Private Non-Profit: 0.98972
  - Public:             0.038005

- Specificity
  - Private For-Profit: 0.992413
  - Private Non-Profit: 0.04792
  - Public:             0.992285


## Map of Treatments across the US

```{r, map code, echo=FALSE, include=FALSE}
facilities_count <- read_csv("~/RStudio/Final Project/Facilities Data/facilities_final.csv")

shapeurl <- "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json"
UnitedStates <- geojson_read(shapeurl, what = "sp")

myLabels <- paste("<strong>", UnitedStates$name, "</strong>", "<br/>", 
                   "Total Number of Facilities: ", facilities_count$Total)
myPopupsTreat <- paste("General Psychotherapy: ", facilities_count$Psychotherapy, "<br/>", "Electroconvulsion Therapy: ", facilities_count$Electroconvulsive, "<br/>", "Substance Abuse Therapy: ", facilities_count$`Substance Abuse`, "<br/>", "Emergency Services: ", facilities_count$Emergency, "<br/>", "Alzheimer's or Dementia Treatment: ", facilities_count$`Alzheimer's or Dementia`, "<br/>", "Eating Disorder Treatment: ", facilities_count$`Eating Disorder`, "<br/>", "PTSD Treatment: ", facilities_count$PTSD, "<br/>", "Private For-Profit: ", facilities_count$`Private For-Profit`, "<br/>", "Private Non-Profit: ", facilities_count$`Private Non-Profit`, "<br/>", "Public: ", facilities_count$Public)


pal_prop <- colorBin("plasma", domain = facilities_count$Proportion)
pal_total <- colorBin("plasma", domain = facilities_count$Total)

Treatment_Map_2019 <- left_join(data.frame(Name = UnitedStates$name), facilities_count, by = c("Name" = "State"))
Treatment_Map_2019 <- leaflet(UnitedStates) %>% addTiles() %>% 
  addPolygons(
    fillColor = pal_prop(facilities_count$Proportion), 
    weight = 2,
    opacity = 1,
    color = "black",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3,
                   color = "black",
                   fillOpacity = 0.7,
                   bringToFront = FALSE),
    label = lapply(myLabels, HTML), 
    popup = myPopupsTreat) %>% 
  addLegend(pal = pal_prop, values = facilities_count$Proportion,
            title = "Facilities to Population (%)", position = "bottomright")
```

```{r, map, echo=FALSE}
Treatment_Map_2019
```
